Presentation Title

### 2017-08-17 thesis/chapters/ch1-intro.tex
Rework ch1 and ch2; fix title

Ironically, also in 2008, Satoshi Nakamoto published the Bitcoin whitepaper~\cite{bitcoin}.
Which, for the first time, gave us a simple banking system in the form of a distributed ledger, also known as blockchain.
It needs no central control but still incorruptible with high probability even if there are malicious parties that aim to undermine the system.

@@

Sadly, as traditional blockchain systems began to gain popularity,
their limitations also became apparent.
Bitcoin has the infamous 7 transactions per second upper bound~\cite{vukolic2015quest}.
This is due to the fact that blocks are fixed to 1 MB,
and are only generated on average every 10 minutes.
Since every Bitcoin transaction is at least 250 bytes, it computes to about 7 transactions per second.
Due to this limitation, it is not uncommon to see a long backlog of about 20,000 transactions\footnote{\url{https://blockchain.info/unconfirmed-transactions/}}.
A few months ago the backlog even reached 100,000, which meant new transactions would take at least 11 hours to be written to the blockchain~\cite{bitcoinbacklog}.
This issue has plagued the Bitcoin community for some time and is the root cause for the block size debate, which some calls it a civil war~\cite{bitcoincivilwar}.
Parties that are for the increase in block size argue that a larger block would improve the transaction rate.
Parties against it argue that it would make mining more centralised because blocks take longer to propagate through the network.
It also requires a hard (non-backwards compatible) fork,
which risks consensus failure and devaluation.
A recent empirical study by Croman et el.~\cite{croman2016scaling} has shown that increasing the block size or the block interval may help.
But given the bandwidth and latency constraints of today's network,
it is not possible to have more than 27 transactions per second.
To achieve transaction rates higher than that, it will lead to unfairness for nodes that do not have a good network connection.
Thus fundamental changes are necessary to run at the scale of centralised payment processors such as Visa,
which is in the order of tens of thousands transactions per second~\cite{visa}.

@@

\section{Contributions}
In this work, we take the advice of Croman et el.~\cite{croman2016scaling} and Vukoli\'{c}~\cite{vukolic2015quest} and rethink the blockchain architecture.
Our primary insight came from observing the differences between how value transfer work in the real world and how they work in early blockchain systems.
Take a restaurant owner, for example,
most of the time the customer is honest and pays the bill.
There is no need for the customer or the owner to report the transaction to any central authority 
because both parties are happy with the transaction.
On the other hand, if the customer leaves without paying the bill,
then the owner would report the incident to some central authority, e.g. the police.
On the contrary, in blockchain systems, every transaction is effectively sent to the miners,
which can be seen as a collective authority.
This consequently leads to limited scalability because every transaction must be validated by the authority even when most of the transactions are legitimate.

Using the aforementioned insight,
we explore an alternative consensus model for blockchain systems where transactions themselves do not reach consensus,
but nevertheless verifiable at a later stage by any node in the network.
We call our protocol \textsf{Checo}, derived from \emph{che}ckpoint \emph{co}nsensus.
Informally, our system works as follows.
Every node stores its own blockchain and every block is one transaction.
We randomly select nodes in every round, the selected ones are called facilitators.
Facilitators reach consensus not on the individual transactions,
but on the state of every chain represented by a single digest, called a checkpoint block.
If a checkpoint block of some node is in consensus, 
then that node can proof to any other node that it holds a set of transactions that compute (form a chain) to the checkpoint block.
This immediately shows that those transactions are tamper-proof.
The key difference from related work is the independence of the transaction validation phase.
Prior systems such a Bitcoin and Ethereum entangled the consensus algorithm with the transaction validation, restricting performance.
By splitting this functionality we achieved horizontal scalability in the general case.
That is, by adding new nodes into the network, the global transaction rate should increase proportionally.

@@

\section{Document structure}
We describe and analyse our idea in the remainder of this work.
We begin with the problem description in~\Cref{ch:problem} which describes the state-of-the-art work on making blockchain systems scalable and our research question.
In~\Cref{ch:model} we give a formal description of our system model and architecture.
Next, we analyse the correctness and performance of our design in~\Cref{ch:analysis}, this is where we present our key theorems.
Implementation and experimental results are discussed in~\Cref{ch:implementation}.
Finally, we conclude in \ref{ch:conclusion} and discuss future work.

>>>>

Ironically, during the midst of the financial crisis, Satoshi Nakamoto published the Bitcoin whitepaper~\cite{bitcoin}.
Which, for the first time, gave us a simple banking system in the form of a distributed ledger, also known as blockchain.
It needs no central control but still incorruptible with high probability even if there are malicious parties that aim to undermine the system.

@@

Sadly, as traditional blockchain systems began to gain popularity,
their limitations also became apparent.
Bitcoin has the infamous 7 transactions per second upper bound~\cite{vukolic2015quest}, 
which is due to the fact that blocks are fixed to 1 MB and are only generated on average every 10 minutes.
Since every Bitcoin transaction is at least 250 bytes, it computes to about 7 transactions per second.
Due to this limitation, it is not uncommon to see a long backlog of about 20,000 transactions\footnote{\url{https://blockchain.info/unconfirmed-transactions/}}.
A few months ago the backlog even reached 100,000, which meant new transactions would take at least 11 hours to be written to the blockchain~\cite{bitcoinbacklog}.
This issue has plagued the Bitcoin community for some time and is the root cause for the block size debate, which some call a civil war~\cite{bitcoincivilwar}.
Parties that are for the increase in block size argue that a larger block would improve the transaction rate.
Parties against it argue that it would make mining more centralised because blocks take longer to propagate through the network.
It also requires a hard (non-backwards compatible) fork,
which risks consensus failure and devaluation.
A recent empirical study by Croman et al.~\cite{croman2016scaling} has shown that increasing the block size or the block interval may help.
But given the bandwidth and latency constraints of today's network,
it is not possible to have more than 27 transactions per second.
To achieve transaction rates higher than that, it will lead to unfairness for nodes that do not have a good network connection.
Thus fundamental changes are necessary to run at the scale of centralised payment processors such as Visa,
which is in the order of tens of thousands transactions per second~\cite{visa}.

@@

\section{Contributions}
In this work, we take the advice of Croman et al.~\cite{croman2016scaling} and Vukoli\'{c}~\cite{vukolic2015quest} and rethink the blockchain architecture.
Our primary insight came from observing the differences between how value transfer work in the real world and how they work in early blockchain systems.
Take a restaurant owner, for example,
most of the time the customer is honest and pays the bill.
There is no need for the customer or the owner to report the transaction to any central authority 
because both parties are happy with the transaction.
On the other hand, if the customer leaves without paying the bill,
then the owner would report the incident to some central authority, e.g. the police.
On the contrary, in blockchain systems, every transaction is effectively sent to the miners,
which can be seen as a collective authority.
This consequently leads to limited scalability because every transaction must be validated by the authority even when most of the transactions are legitimate.

Using the aforementioned insight,
we explore an alternative consensus model for blockchain systems where transactions themselves do not reach consensus,
but nevertheless verifiable at a later stage by any node in the network.
We call our protocol \textsc{Checo}, derived from \emph{che}ckpoint \emph{co}nsensus.
Informally, our system works as follows.
Every node stores its own blockchain and every block is one transaction.
We randomly select nodes in every round, the selected ones are called facilitators.
Facilitators reach consensus not on the individual transactions,
but on the state of every chain represented by a single digest, called a checkpoint block.
If a checkpoint block of some node is in consensus, 
then that node can prove to any other node that it holds a set of transactions that compute (form a chain) to the checkpoint block.
This immediately shows that those transactions are tamper-proof.
The key difference from related work is the independence of the transaction validation phase.
Prior systems such a Bitcoin and Ethereum entangled the consensus algorithm with the transaction validation, restricting performance.
By splitting this functionality we achieved horizontal scalability in the general case.
That is, by adding new nodes into the network, the global transaction rate should increase proportionally,
regardless of the choice of system parameters.

@@

\section{Document structure}
We begin with the problem description in~\Cref{ch:problem},
it first describes the state-of-the-art in scalable blockchain systems and their shortcomings,
which leads to our research question.
In~\Cref{ch:model} we give a formal description of our system model and architecture.
Next, we analyse the correctness and performance of our design in~\Cref{ch:analysis}, this is where we present our key theorems.
Implementation and experimental results are discussed in~\Cref{ch:implementation}.
Finally, we conclude in~\Cref{ch:conclusion} and discuss future work.

### 2017-08-17 thesis/chapters/ch2-problem.tex
Rework ch1 and ch2; fix title

\subsection{Early blockchain systems}
This category represents the baseline,
which are systems with a probabilistic consensus algorithm.
That is, transactions never reach consensus with a probability of 1.
The typical examples are proof-of-work based systems such as Bitcoin, Ethereum and other Altcoins.
In Bitcoin, nodes compete in solving puzzles.
The first node solves the puzzle has the privilege to create a new block.
The level of consensus of a block containing many transactions is determined by how deep it is in the Bitcoin blockchain,
also called the number of confirmation.
The probability of a block being orphaned drops exponentially as the depth increases~\cite{bitcoin}.
Nevertheless, the probability of the highest block being orphaned is non-negligible.

@@

This category of systems uses traditional Byzantine consensus algorithms such as PBFT~\cite{castro1999practical}.
In essence, they contain a fixed set of nodes (sometimes called validating peer) that run a Byzantine consensus algorithm to decide on new blocks.
This is often used in permissioned system where the validators must be predetermined,
for example, Hyperledger Fabric~\cite{cachin2016architecture}.

@@

We saw in~\Cref{sec:taxonomy} that the majority of the systems cannot achieve horizontal scalability.
To the best of our knowledge, only two systems---Elastico~\cite{luu2016elastico} and OmniLedger~\cite{kokoris2017omniledger}---has experimentally demonstrated horizontal scalability.
However, these techniques rely on sharding, thus the performance highly depends on transaction characteristics and the shard size parameter because inter-shard transactions are slow.
Furthermore, all the systems described in~\Cref{sec:taxonomy} are complete blockchain systems designed for a concrete application.
We wish to design a blockchain consensus protocol that can be used in many applications.
For example recording uploads and downloads between peers in Tribler~\cite{pouwelse2008tribler, multichain} and decentralised market\footnote{\url{https://github.com/Tribler/tribler/blob/devel/Tribler/community/tradechain/__init__.py}}.
Hence our research question is the following.
\begin{displayquote}
\emph{Is it possible to design a blockchain consensus protocol that is fault-tolerant, scalable and can reach global consensus?}
\end{displayquote}

@@

In the remainder of this section, we expand on our research question and visit each of our requirement in detail.
At the end of this section, we describe some issues that we do not consider as a part of our design,
which is the result of not focusing on any specific application.

\subsection{Global consensus}

% Early blockchain systems use PoW as the consensus algorithm,
% but consensus algorithm and especially Byzantine\footnote{Byzantine means participants can experience arbitrary failure, more on this in~\Cref{sec:acs-background}.}
% consensus algorithms have existece since the 80s~\cite{pease1980reaching, lamport1982byzantine}.
% The Practical Byzantine Fault Tolerance (PBFT) algorithm introduced in 1999 by Castro and Liskov~\cite{castro1999practical} is widely regarded as the first high-performance consensus algorithm.
% The authors show that running PBFT in Network File System (NFS) was only 3\% slower than the standard unreplicated NFS.
% Since then, research in this area picked up pace and many more algorithms are introduced, for example Q/U~\cite{abd2005fault} and Zyzzyva~\cite{kotla2007zyzzyva}.
% With the introduction of Bitcoin's probabilistic consensus model, the field experienced a second renaissance.
% These include proof-of-stake protocols~\cite{bentov2016snow, micali2016algorand}, as well as protocols optimised for blockchain~\cite{liu2016xft, miller2016honey}.

The first objective is to reach consensus on a global \emph{state} by building on top of decades of Byzantine consensus protocol research~\cite{pease1980reaching, lamport1982byzantine, castro1999practical, miller2016honey}.
Having consensus is essential in blockchain systems.
It stops many types of malicious activities because agreed state must have the consent of the honest nodes in the network.
We emphasise the word state because in contrary to traditional blockchains,
it does not necessarily mean the set of all transactions that have ever happened.
The state can be a verifiable representation of all the transactions.
As an extreme example, a single digest of the complete history can represent the state.
It is even verifiable as long as there are nodes that store the history.

% Our approach is to combine the vast literature in Byzantine fault tolerance research with blockchain technology,
% and introduce a new consensus model to record the global state.

\subsection{Security and fault-tolerance}

Our system should be unaffected in the presence of powerful adversaries.
In particular, adversaries are Byzantine meaning that they can have arbitrary behaviour.
Thus anything is possible from simply omitting messages to colluding with each other in order to undermine the whole system.
This aspect usually comes for free when using a Byzantine fault-tolerant consensus algorithm.

There is another aspect of fault tolerance that is transaction verifiability.
All transactions must be eventually verifiable to maintain system integrity.
Further, the validation result must be consistent.
For example, if two different honest nodes attempts verify the same transaction,
it cannot be the case that one node thinks that the transaction is valid but the other thinks it is invalid.

% Recall that it is not neccessary to reach consensus on all the transactions in order to record the state.
% But when using some digest (or a set of digets) to represent the state, it is also important for the transactions to be verifiable.
% Otherwise adversaries could simply store invalid transactions in the state and they would never be detected.

\subsection{Performance and scalability}

The last but not least, the final objective is horizontal scalability.
We wish to see the performance (global transaction rate) increase as more nodes join the network.
BitTorrent~\cite{cohen2003incentives} is an example of such a system,
where peers interact with each other to exchange files without any global bottleneck.

Early blockchain systems may run in a network of thousands of nodes, but the performance does not scale.
On the other hand, as we will see in~\Cref{sec:acs-background}, Byzantine consensus algorithms perform well but only when in a small network.
In this work, our goal is to overcome the limitations of early blockchain systems and Byzantine consensus algorithms to create a horizontally scalable blockchain.

\subsection{Limitations}
Our work aims to significantly lower the transaction bound.
However, the price to pay is that the Sybil attack~\cite{douceur2002sybil} prevention mechanism moves from the blockchain consensus protocol to the application layer.
For example, social network based Sybil defence mechanisms use graph structure of real-world relationships~\cite{yu2006sybilguard}.
Online marketplaces such as Amazon use the rating of buyer and sellers.
Since our system is application neutral, we do not provide an explicit Sybil defence mechanism.
On the other hand, our system also has no restrictions on the Sybil defence technique
and applications built on top of our system can use the best mechanism for their purpose.

>>>>

\subsection{Early blockchain systems}
This category represents the baseline,
which are systems with a probabilistic consensus algorithm.
That is, transactions never reach consensus with a probability of 1.
Given sufficient computational power, adversaries are able to undo any transaction.
The typical examples are proof-of-work based systems such as Bitcoin, Ethereum and other Altcoins.
In Bitcoin, nodes compete in solving puzzles.
The first node solves the puzzle has the privilege to create a new block.
The level of consensus of a block containing many transactions is determined by how deep it is in the Bitcoin blockchain,
also called the number of confirmation.
The probability of a block being orphaned drops exponentially as the depth increases~\cite{bitcoin}.
Nevertheless, the probability of the highest block being orphaned is non-negligible.

@@

This category of systems uses traditional Byzantine consensus algorithms such as PBFT~\cite{castro1999practical}.
In essence, they contain a fixed set of nodes (sometimes called validating peer) that run a Byzantine consensus algorithm to decide on new blocks.
This technique is often used in permissioned system where the validators must be predetermined.
For example, Hyperledger Fabric~\cite{cachin2016architecture}.

@@

We saw in~\Cref{sec:taxonomy} that the majority of the systems cannot achieve horizontal scalability.
To the best of our knowledge, only two systems---Elastico~\cite{luu2016elastico} and OmniLedger~\cite{kokoris2017omniledger}---has experimentally demonstrated horizontal scalability.
However, these techniques rely on sharding, thus the performance highly depends on transaction characteristics and the shard size parameter because inter-shard transactions are slow.
Furthermore, all the systems described in~\Cref{sec:taxonomy} are complete blockchain systems designed for a concrete application.
We wish to design a blockchain consensus protocol that can be used in many applications.
For example, the accounting of detailed internet traffic in Tribler~\cite{pouwelse2008tribler, multichain} and decentralised market\footnote{\url{https://github.com/Tribler/tribler/blob/devel/Tribler/community/tradechain/__init__.py}}.
Hence our research question is the following.
\begin{displayquote}
\emph{How can we design a blockchain consensus protocol that is fault-tolerant,
horizontally scalable,
and able to reach global consensus?}
\end{displayquote}

@@

In the remainder of this section, we expand on our research question and visit each of our requirements in detail.
At the end of this section, we describe some issues that we do not consider as a part of our design,
which is the result of not focusing on any specific application.

\subsection*{Global consensus}
% WHY WE NEED GLOBAL CONSENSUS?
Having consensus is essential in blockchain systems.
It stops many types of malicious activities because the agreed state, or a representation of it, must have the consent of the honest nodes in the network.
We emphasise on the notion of a representation of the state because in contrary to traditional blockchains,
it does not necessarily mean the set of all transactions that have ever happened.
The state can be a small, but verifiable representation of all the transactions.
As an extreme example, a single digest of the complete history can represent the state.
It is even verifiable as long as there are nodes that store the history.

\subsection*{Fault-tolerance}
% WHY WE NEED FAULT-TOLERANCE
Similar to many blockchain systems,
our system should be unaffected in the presence of powerful adversaries.
In particular, adversaries are Byzantine meaning that they can have arbitrary behaviour.
Thus anything is possible from simply omitting messages to colluding with each other in order to undermine the whole system.

There is another aspect of fault tolerance that is transaction verifiability.
All transactions must be eventually verifiable to maintain system integrity.
Further, the validation result must be consistent.
For example, if two different honest nodes attempt to verify the same transaction,
it cannot be the case that one node thinks that the transaction is valid but the other thinks it is invalid.

\subsection*{Horizontal scalability}

To enable ubiquitous use, we demand horizontal scalability.
That is, the global transaction rate should increase as more nodes join the network.
BitTorrent~\cite{cohen2003incentives} is an example of such a system,
where peers interact with each other to exchange files without any global bottleneck.
% Early blockchain systems may run in a network of thousands of nodes, but the performance does not scale.
% On the other hand, as we will see in~\Cref{sec:acs-background}, Byzantine consensus algorithms perform well but only when in a small network.
% In this work, our goal is to overcome the limitations of early blockchain systems and Byzantine consensus algorithms to create a horizontally scalable blockchain.

\subsection*{Limitations}
Since our consensus protocol is application neutral, it does not attempt to prevent the Sybil attack~\cite{douceur2002sybil}.
A good defence mechanism requires application specific information.
For example, a social network based Sybil defence mechanisms use graph structure of real-world relationships~\cite{yu2006sybilguard}.
Online marketplaces such as Amazon use the rating of buyer and sellers.
% Since our system is application neutral, we do not provide an explicit Sybil defence mechanism.
On the other hand, our system also has no restrictions on the Sybil defence mechanism
and applications built on top of our system can use the best mechanism for their purpose.

### 2017-08-17 thesis/formalities/abstract.tex
Rework ch1 and ch2; fix title

\noindent \textbf{Author}\\
\begin{tabular}{l}
Kelong Cong\\
\\
\end{tabular}\\
\noindent \textbf{Title}\\
\begin{tabular}{l}
Blockchain Consensus Protocol with Horizontal Scalability\\
\\
\end{tabular}\\
\noindent \textbf{MSc presentation}\\
\begin{tabular}{l}
% <MM> DD, YYYY (like \today)
31st August 2017\\
\\
\end{tabular}

@@

\begin{abstract}
% what's the problem?
Blockchain systems have the potential to decentralise many traditionally centralised systems.
However, scalability remains a key challenge.
Without a horizontally scalable solution, blockchain systems remain unsuitable for ubiquitous use.
% what's the solution?
We design a novel blockchain system called \textsc{Checo}.
Each node in our system maintains individual hash chains,
which only stores transactions that the node is involved in.
A consensus is reached on special blocks called checkpoint blocks rather than on all the transactions.
Checkpoint blocks are effectively a hash pointer to the individual hash chains,
thus a single checkpoint block may represent a large set of transaction blocks.
The consensus protocol does not imply transaction validity.
Hence we include a validation protocol which allows nodes to verify that transactions are correctly recorded,
which indirectly achieves consensus on transactions.
% how good is the solution?
We implement a prototype and evaluate it experimentally.
Our results show a strong indication of horizontal scalability even in the worst case,
where every transaction is made with a randomly selected node.
For 1200 nodes, we were able to perform almost 5000 transactions per second,
orders of magnitude higher than the 7 transactions per second limit of Bitcoin.
\end{abstract}

>>>>

\noindent \textbf{Author}\\
\begin{tabular}{l}
Kelong Cong\\
\\
\end{tabular}\\
\noindent \textbf{Title}\\
\begin{tabular}{l}
A Blockchain Consensus Protocol With Horizontal Scalability\\
\\
\end{tabular}\\
\noindent \textbf{MSc presentation}\\
\begin{tabular}{l}
% <MM> DD, YYYY (like \today)
31st August 2017\\
\\
\end{tabular}

@@

\begin{abstract}
% what's the problem?
Blockchain systems have the potential to decentralise many traditionally centralised systems.
However, scalability remains a key challenge.
Without a horizontally scalable solution, blockchain systems remain unsuitable for ubiquitous use.
% what's the solution?
We design a novel blockchain system called \textsc{Checo}.
Each node in our system maintains individual hash chains,
which only stores transactions that the node is involved in.
A consensus is reached on special blocks called checkpoint blocks rather than on all the transactions.
Checkpoint blocks are effectively a hash pointer to the individual hash chains,
thus a single checkpoint block may represent a large set of transaction blocks.
The consensus protocol does not imply transaction validity.
Hence we include a validation protocol which allows nodes to verify that transactions are correctly recorded,
which indirectly achieves consensus on transactions.
% how good is the solution?
We implement a prototype and evaluate it experimentally.
Our results show a strong indication of horizontal scalability even in the worst case,
where every transaction is made with a randomly selected node.
\end{abstract}

### 2017-08-17 thesis/formalities/preface.tex
Rework ch1 and ch2; fix title

\chapter*{Preface}
\addcontentsline{toc}{chapter}{Preface}
Since early 2013, when I mined my first Bitcoin, 
I have been interested in decentralising the traditionally centralised systems.
Following the Snowden revelations, the interest became a desire that has stayed with me ever since.
Today, many blockchain systems are mature, but still not yet ready for ubiquitous use due to the scalability limitation.
Many of us still do not use Bitcoin to purchase everyday goods, online or offline.
This limitation, as well as my personal ambitions, led me to this thesis,
with the goal of creating a truly horizontally scalable solution.

>>>>

\chapter*{Preface}
\addcontentsline{toc}{chapter}{Preface}
Since early 2013, the time when I mined my first Bitcoin, 
I have been interested in decentralising the traditionally centralised systems.
Following the Snowden revelations, the interest became a desire that has stayed with me ever since.
Today, many blockchain systems are mature, but still not yet ready for ubiquitous use due to the scalability limitation.
Many of us still do not use Bitcoin to purchase everyday goods, online or offline.
This limitation, as well as my personal ambitions, led me to this thesis,
with the goal of creating a truly horizontally scalable solution.
